---
title: "Singular Value Decomposition"
output: html_document
---


```{r}
library(tm)
library(tokenizers)
library(stringr)

stories <- scan("gutenberg short stories-2.txt", what = "character", allowEscapes = TRUE)
all_words <- paste(stories[1:length(stories)], collapse = " ")
```

Description of Data
- 13 short stories by 13 authors 
- Check the length of each short story by words, sentences, and paragraph

Preprocessing

The tm package created by Ingo Feinerer and Kurt Hornik (2020) makes it possible to swiftly create a document term matrix.

```{r}
#edit abbreviations/text formatting
all_words <- gsub("Mrs.", "Mrs", all_words)
all_words <- gsub("Mr.", "Mr", all_words)
all_words <- gsub("said", ",said", all_words)
all_words <- gsub("says", ",says", all_words)
all_words <- gsub("Says", ",says", all_words)


sentences <- tokenize_sentences(all_words) # creates list


# make a table indexing each sentence
sentence_index <- data.frame(sentences = sentences ,
                         index = 1:length(sentences )) #to sort by story 


var_names <- sentence_index$c..This.thing.is.so.rare.in.all.its.circumstances..and.on.so.good.authority..that.my.reading.and.conversation.have.not.given.me.anything.like.it....

#convert to vector
for(i in var_names) {
    assign(i, sentences[[i]])
}

# sentences[[716]] # fails to retrieve  
# unlist(x)
(abbrev <- str_extract(var_names, "^(?=.*\\.)[A-Za-z.]{4}$")) #retrieve abbreviations
# grep("^(?=.*[.])[a-zA-Z.]{3}$", x, value = TRUE, er)


#preparing sentence documents
var_names <- removePunctuation(var_names)
var_names <- removeNumbers(var_names)
var_names <- tolower(var_names)

#checking for low/high frequency words
all_words_post <- as.vector(var_names) #return to single string

X <- removeWords(all_words_post,stopwords(kind = "en"))
class(X)
X_vector <- length(unique(as.vector(X)))

unique_words <- unique(all_words_post) 
length(unique_words)
class(unique_words)

x <- removeWords(X_vector, unique_words)

x1 <- remove(X_vector, unique_words)


docs <- var_names
sentence_corpus <- VCorpus(VectorSource(docs))
sentence_corpus <- tm_map(sentence_corpus, removeWords, stopwords('english'))
class(sentence_corpus)


sentence_dtm <- TermDocumentMatrix(sentence_corpus)
(sentence_dtm <- as.matrix(sentence_dtm))

#weight matrix

#count words
# sapply(strsplit(all_words_post," "), length)
# sapply(strsplit(unique_words," "), length)

```

## Characteristics of SVD

* Saves most important data
* Each column vector of the resultant matrix represents each component found through SVD. The components are ordered by amount of variation, starting from greatest to least. 
* Reduces data through the use of eigenvalues representing primary components
* The SVD equation in words : The original matrix is equal to an orthogonal matrix U, multipied by a diagonal matrix containing eigenvalues multiplied by the inverse matrix of U, named V. 

Example of SVD:
```{r}
a <- c(2,0,8,6,0,
        1,6,0,1,7,
        5,0,7,4,0,
        7,0,8,5,0,
        0,10,0,0,7)
A <- matrix(a,byrow=TRUE, ncol=5)
At <- t(A)

#Step 1 - multiply A and A transposed
U<-A%*%At

eigen(U)

#Step 2 - multiply by vertical variable matrix* and use system of equations to solve for lambda.

#Step 3 - from the paired eigenvalues(lambda), create a new matrix arranged by column from largest eigenvalues to smallest. 

V <- At%*%A
```
$$
A\vec{v} = \lambda\vec{v}
$$
From U and V, find lambda using above equation. In this example, the determinant is set to 0 in order to use system of equations and solve for lambda.**

Gram-Schmidt orthonormalization process is applied to create orthogonal matrices: 
To normalize a vector, the following equation applies:
$$
\hat{v}=\frac{\vec{v}}{|\vec{v}|}
$$
$$
|\vec{v}| 
$$ is calculated by dividing by the magnitude (or length?) of the vector. The quotient is the unit vector

Example 2 of SVD (Landauer and Dumais, pg. 10, Figure 1)

$$
{X } = {W}{S}{P}'
$$

Reproducing Landauer and Dumais example. 
```{r}
freq<-c(1,0,0,1,0,0,0,0,0,
        1,0,1,0,0,0,0,0,0,
        1,1,0,0,0,0,0,0,0,
        0,1,1,0,1,0,0,0,0,
        0,1,1,2,0,0,0,0,0,
        0,1,0,0,1,0,0,0,0,
        0,1,0,0,1,0,0,0,0,
        0,0,1,1,0,0,0,0,0,
        0,1,0,0,0,0,0,0,1,
        0,0,0,0,0,1,1,1,0,
        0,0,0,0,0,0,1,1,1,
        0,0,0,0,0,0,0,1,1)
A <- matrix(freq, ncol=9, nrow=12, byrow=TRUE)
A
rownames(A)<-c("1human", "2interface", "3computer", "4user", "5system", "6response", "7time", "8EPS", "9survey", "10trees", "11graph", "12minors")

cor(A["1human",],A["12minors",])
cor(A["1human",],A["4user",])

W <- A%*%t(A)

e <- eigen(W)
W <- e$vectors
W <- W*-1

ex <- c(-1,1,
        -1,3,
        -3,1,
        -3,3)
ex <- matrix(ex, ncol=2, byrow=TRUE)
t(ex)

```



References:
Kirk Baker, SVD Tutorial, 2005 
Landauer and Dumais, 1998

